Are we living in a depressive society?

### 2024-04-15 outputs/paper/paper.qmd
commit

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-alonetrend-graph
#| fig-cap: "Adults who usually or always feel lonely"

# Define the path to the data file and load it
social_data <- read_parquet(file = here::here("outputs/data/Social_Connection.parquet"))

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Age" subgroup
social_age_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Age")

# Plot the stacked bar chart
graph_age <- ggplot(social_age_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Age",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Sex" subgroup
social_sex_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Sex")

# Plot the stacked bar chart
graph_sex <- ggplot(social_sex_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Sex",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Race/Hispanic ethnicity" subgroup
social_race_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Race/Hispanic ethnicity")

# Plot the stacked bar chart
graph_race <- ggplot(social_race_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Race",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Education" subgroup
social_educ_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Education")

# Plot the stacked bar chart
graph_educ <- ggplot(social_educ_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Education",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Arrange graphs in single pane
grid.arrange(graph_age, graph_sex, graph_race, graph_educ, nrow = 2, ncol = 2)

@@

## Model set-up

@@

### Discussion of Results
@fig-ageandtrend-graph indicates that while all age groups experienced fluctuations in depressive disorder rates, younger adults consistently reported higher rates. This could be due to a variety of factors including economic instability, career pressures, or social isolation, particularly accentuated by the pandemic. Trends of older adults, age 50 or more, might reflect accumulating stresses or health-related issues as they age, especially during challenging times like the pandemic. 
The peaks and valleys in the @fig-ageandtrend-graph correlate with the unfolding of global events, particularly the COVID-19 pandemic, which has been a significant stressor impacting mental health across all demographics. The data shows how these impacts vary significantly with age, suggesting that age-specific interventions might be necessary to address the mental health needs of different populations effectively.

@@

____
### Implications
The data underscores the need for targeted mental health services and interventions that cater specifically to the unique needs of different age groups. Understanding these trends helps in planning public health strategies and allocating resources more effectively to manage and mitigate depressive disorders among the population.

@@

The depressive disorder rates among different educational groups emphasize the need for policy interventions that consider the intersection of education, economic stability, and mental health. Tailoring mental health services to meet the needs of people across different educational backgrounds can lead to more effective prevention and treatment strategies, ultimately reducing the overall burden of depressive disorders in society.
____

>>>>

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(gridExtra)
library(readr)

# Load the data
data <- read_parquet(file = here::here("outputs/data/Social_Connection.parquet"))

# Filter data for the specific indicator
lonely_data <- data %>%
  filter(Indicator == "Adults who usually or always feel lonely")

# Function to create a plot for a given group with specified y-axis limits
create_plot <- function(data, group_name, ymax) {
  data_group <- data %>%
    filter(Group == group_name) %>%
    group_by(Time.Period.Label, Subgroup) %>%
    summarise(Value = sum(Value), .groups = "drop") %>%
    arrange(Time.Period.Label, Subgroup)

  # Add cumulative sum of Value for label positioning
  data_group <- data_group %>%
    group_by(Time.Period.Label) %>%
    mutate(label_pos = cumsum(Value) - 0.5 * Value,
           segment_size = Value)

  # Create the plot
  plot <- ggplot(data_group, aes(x = Time.Period.Label, y = Value, fill = Subgroup)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(y = label_pos, label = ifelse(segment_size > 5, Subgroup, "")), 
              position = position_stack(vjust = 0.5), 
              color = "white", family = "Times", check_overlap = TRUE, 
              size = ifelse(data_group$segment_size > 5, 3, 0)) +
    labs(title = paste(group_name),
         y = "Symptom rates (%)",
         x = NULL) +
    ylim(0, ymax) +
    theme_minimal() +
    theme(text = element_text(family = "Times"),
          axis.text.x = element_text(hjust = 1, size = 10),
          plot.title = element_text(size = 7),
          legend.position = "none")

  return(plot)
}

# Create plots for each group with specific y-axis limits
plot_age <- create_plot(lonely_data, "By Age", 100)
plot_sex <- create_plot(lonely_data, "By Sex", 30) 
plot_education <- create_plot(lonely_data, "By Education", 70)
plot_race <- create_plot(lonely_data, "By Race/Hispanic ethnicity", 70)

# Arrange plots in a 2 x 2 grid
grid_plots <- grid.arrange(plot_age, plot_sex, plot_education, plot_race, ncol = 2, nrow = 2)

# Print the arranged plots
print(grid_plots)

@@

## Model set-up
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma)
\end{align}

@@

@fig-ageandtrend-graph indicates that while all age groups experienced fluctuations in depressive disorder rates, younger adults consistently reported higher rates. This could be due to a variety of factors including economic instability, career pressures, or social isolation, particularly accentuated by the pandemic. Trends of older adults, age 50 or more, might reflect accumulating stresses or health-related issues as they age, especially during challenging times like the pandemic. 
The peaks and valleys in the @fig-ageandtrend-graph correlate with the unfolding of global events, particularly the COVID-19 pandemic, which has been a significant stressor impacting mental health across all demographics. The data shows how these impacts vary significantly with age, suggesting that age-specific interventions might be necessary to address the mental health needs of different populations effectively.

@@

The data underscores the need for targeted mental health services and interventions that cater specifically to the unique needs of different age groups. Understanding these trends helps in planning public health strategies and allocating resources more effectively to manage and mitigate depressive disorders among the population.

@@

The depressive disorder rates among different educational groups emphasize the need for policy interventions that consider the intersection of education, economic stability, and mental health. Tailoring mental health services to meet the needs of people across different educational backgrounds can lead to more effective prevention and treatment strategies, ultimately reducing the overall burden of depressive disorders in society.
